Backpropagating Gravitational wave events

### 2022-05-23 src/ms.tex
citation explosion

\section{Introduction}
As the detection rate of gravitation waves (GWs) from merging
double-compact-object (DCO) binaries increases with the sensitivity of the
ground-based GW detector network
\citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021},
we are beginning to constrain the astrophysical processes which shape the
evolution of GW progenitor populations. One of the most common ways to study
progenitor populations of GW mergers is through population synthesis simulations
for a host of formation environments including isolated binary stars, triple
systems, globular clusters, nuclear star clusters, AGN disks, and more
\citep{Mandel2022}. \wf{Moar citations here} Traditionally in these studies, the
population is studied through a Monte-Carlo approach.  Using theoretically
motivated distributions of initial parameters (age, metallicity, masses, orbital
separations, eccentricities, etc), initial populations are evolved with fixed
astrophysical assumptions (population synthesis parameter settings) to produce a
synthetic catalog of mergers which can be compared to observations from
GW detectors.

@@

While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
One of the most common skepticisms of population synthesis studies is that \kb{they} employ simulation sets which often 
cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
On top of this, these values are often not calibrated to data.
Even with recent developments in using emulators to speed up the simulation process, training the emulators 
still requires a significant amount of computation to cover a wide range of uncertainty.
As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
run a grid which varies combinations of uncertain physics.
For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
this would still require $1024$ simulations.
Because population synthesis simulations require hundreds to thousands of cpu hours,
training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.

>>>>

\section{Introduction}
As the detection rate of gravitation waves (GWs) from merging
double-compact-object (DCO) binaries increases with the sensitivity of the
ground-based GW detector network
\citep{Abbott2018,LIGOScientificCollaboration2015,Buikema2020,Tse2019,Acernese2015,Acernese2019,Aso2013,Akutsu2021},
we are beginning to constrain the astrophysical processes which shape the
evolution of GW progenitor populations. One of the most common ways to study
progenitor populations of GW mergers is through population synthesis simulations
for a host of formation environments. In the case of isolated binary evolution, merging 
double compact objects can form from massive binary stars through the
standard stable mass transfer or common envelope channels 
\citep[e.g.][]{Belczynski2002, Dominik2012, Belczynski2016, Zevin2020, Bavera2021, Broekgaarden2021, VanSon2022} 
as well as through chemically homogenous evolution \citep{Mandel2016, Marchant2016, deMink2016} 
or with population III stars \citep[e.g.][]{Belczynski2004, Kinugawa2014, Inayoshi2016, Inayoshi2017, Tanikawa2021, Tanikawa2022}.
Merging DCO binaries can also originate from a wide variety of dynamically active environments including 
triple (or higher multiple) systems \citep[e.g. ][]{Antonini2017, Silsbee2017, Fragione2019, VignaGomez2021}, 
as well as young stellar clusters \citep[e.g.][]{Ziosi2014, Banerjee2017, DiCarlo2020, Chattopadhyay2022}, 
globular clusters \citep[e.g.][]{PortegiesZwart2000, OLeary2006, Downing2010, Samsing2014, Rodriguez2015, Rodriguez2016, Askar2017, Rodriguez2019}, 
nuclear star clusters \citep[][]{Miller2009, Antonini2016}, or a mix of all three \citep[e.g.][]{Mapelli2022}.
Finally, more exotic channels like active galactic nuclei which combine both gravitational and gas interactions 
\citep[e.g.][]{McKernan2018, McKernan2020, Secunda2020, Ford2021} or primordial black hole formation \textbf{KB add these citations} can lead 
to merging DCOs observable with a ground-based detector network. For a discussion of the relative rates of each 
formation envrionment and channels within, see \citet{Mandel2022}. Traditionally in these studies, the
population is studied through a Monte-Carlo approach.  Using theoretically
motivated distributions of initial parameters (age, metallicity, masses, orbital
separations, eccentricities, etc), initial populations are evolved with fixed
astrophysical assumptions (population synthesis parameter settings) to produce a
synthetic catalog of mergers which can be compared to observations from
GW detectors.

@@

While forward modelling entire populations has been the canonical pathway to compare simulations and GW observations,
it is not necessarily the most efficient way to constrain the underlying physical processes, or ``hyperparameters", which drive the evolution of the population.
One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often 
cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
On top of this, these values are often not calibrated to data.
Even with recent developments in using emulators to speed up the simulation process, training the emulators 
still requires a significant amount of computation to cover a wide range of uncertainty.
As an example, to emulate a model with $10$ parameters, the simplest way to construct a training set of simulations is to 
run a grid which varies combinations of uncertain physics.
For $10$ uncertain physical processes, if we consider $2$ variations for each physical process 
this would still require $1024$ simulations.
Because population synthesis simulations require hundreds to thousands of cpu hours,
training an emulator which spans a high dimensional uncertainty space remains an impossibility at present.
